---
output:
  github_document:
    html_preview: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# mvcapaCor
An `R` package for detecting (collective and point) anomalies or changepoints in cross-correlated data.
It also contains code to reproduce the simulation study in "Scalable changepoint and anaomly detection in cross-correlated data with an application to condition monitoring".

## Overview
Functionality:

* Function for running the CAPA-CC algorithm on data, as well as CPT-CC for a single change.
* Functions for visualising the detected collective and point anomalies.
* Functions for estimating a precision matrix restricted to a given adjacency matrix by the GLASSO method.
* Functions for generating test data and running the simulation study in "Scalable changepoint and anaomly detection in cross-correlated data with an application to condition monitoring".

## Installation

You can install tpcaMonitoring from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("Tveten/mvcapaCor")
```

## Exported and documented functions
For more information, see the documentation of the functions below inside R.

* mvcapa_cor
* estimate_precision_mat
* plot_capa

## Example

```{r example, message = FALSE, warning = FALSE}
library(mvcapaCor)

# Generate the test sets used:
x <- simulate_cor(200, 4)
```
